Deliminative verbal reduplication in Mandarin Chinese

### 2025-06-17 replies-reviews2.tex

%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%
\noindent
\textbf{Answer:} This passage is removed.

\item p. 15, lines 25--28: ``\textit{Le} is compatible with the reduplication, because its dynamicity can relate to not
only the termination or instantiation of an event (a point of change), but also the process of the
situation, just like that of the reduplication (see Section 2.3.1)''. This passage is unclear. Please revise
it for clarity.

%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%
\noindent
\textbf{Answer:} This passage is revised as follows 
and moved before example (27) ((28) in the previous version of the draft) to directly follow the passage before: 
``As shown in Section 2.3.1, reduplication can also express
 dynamicity of both a time point and a time period, 
 just like \obj{le} illustrated here. 
 Therefore, \obj{le} is compatible with reduplication.''

### 2025-06-17 chapters/ch02-Phenomenon.tex
check main text before sec 2.4

\ea
  \ea\label{ex:le-dyn-look}
    \gll \objex{Wu2} \objex{Xu4mang2} \objex{kan4-le} \objex{zuo4-an4} \objex{shi2} \objex{liu2xia4} de \objex{jiao3yin1} \ldots\\ 
    Wu Xumang look-\textsc{pfv} commit-crime when leave \textsc{de} footprint\\ \jambox{\citep[158]{XiaoMcEnery2004}}
    \glt `Wu Xumang looked at the footprint left when the crime was committed.'
  \ex\label{ex:redup-dyn-look}
    \gll  \objex{Wu2} \objex{Xu4mang2} \objex{kan4-le-kan4} \objex{zuo4-an4} \objex{shi2} \objex{liu2xia4} de \objex{jiao3yin1} \ldots\\
    Wu Xumang look-\textsc{pfv}-look commit-crime when leave \textsc{de} footprint\\ \jambox{\citep[158]{XiaoMcEnery2004}}
    \glt `Wu Xumang looked a little bit at the footprint left when the crime was committed.'
  \z
\z

@@

As a reviewer points out, \citet[Sec.\,3.3]{Tham2013} considers  \obj{bing4} `be sick' to be a basic change of state (COS) verb, i.e. `become sick'.
In contrast, she uses \obj{xi3huan1} `like' and \obj{xiang1xin4} `believe' as examples of stative verbs.
\citet[680]{PeckEtAl2013} also list \obj{xi3huan1} `like' and \obj{xiang1xin4} `believe' as stative verbs.
For these two verbs, examples such as (\ref{ex:xihuan}) and (\ref{ex:xiangxin}) can be found.
\ea\label{ex:xihuan}%这车好型啊！！！。怎么了，我就喜欢喜欢 不行么~ (BCC)
\gll \objex{zhe4} \objex{che1} \objex{hao3} \objex{xing2} \objex{a}! \objex{zen3me} \objex{le}, \objex{wo3} \objex{jiu4} \obj{xi3huan1-xi3huan4} \objex{bu4xing2} \objex{me}?\\
this car so cool \textsc{ptc} what \textsc{ptc} I just like-like cannot \textsc{ptc}\\\jambox{(BCC)}
\glt `This car is so cool! What? I can't just like it for a little bit?'

\ex\label{ex:xiangxin}%谁让我来相信相信 (CCL)
\gll \objex{shei2} \objex{rang4} \objex{wo3} \objex{lai2} \objex{xiang1xin4-xiang1xin4}\\
who let I come believe-believe\\\jambox{(CCL)}
\glt `Who lets me believe him/her?'
\z

@@

\ea
  \ea\label{ex:cry}
    \gll \objex{wei4ci3}, \objex{Deng4} \objex{Li4jun1} \objex{shang1xin1} \objex{de} \objex{ku1-le} \objex{san1} \objex{tian1}.\\
    for.this Deng Lijun sadly \textsc{de} cry-\textsc{pfv} three day\\ \jambox{\citep[129]{XiaoMcEnery2004}}
    \glt `For this reason, Deng Lijun cried sadly for three days.'
  \ex\label{ex:know}
    \gll \objex{ta1} \objex{zhi1dao4-le} \objex{zhe4} \objex{shi4} de \objex{nei4-qing2}.\\
    he know-\textsc{pfv} this matter \textsc{de} inside-information\\ \jambox{\citep[130]{XiaoMcEnery2004}}
    \glt `He got to know the inside information on this matter.'
  \z
\z
%\obj{Le} is compatible with the reduplication, because its dynamicity can relate to not only the termination or instantiation of an event (a point of change), but also the process of the situation, just like that of the reduplication (see \sectref{sec:core-sem}).

In comparison, the experiential aspect marker \obj{guo4} cannot co\hyp{}occur with a reduplicated verb, 
because its dynamicity attributes to an ``experiential change'' \citep[148]{XiaoMcEnery2004}, 
namely that a situation has been experienced historically and that ``the final state of the situation no longer obtains'' at the reference time \citep[144]{XiaoMcEnery2004}. 
Compare (\ref{ex:guo-dyn}) and (\ref{ex:le-dyn}), \obj{guo4} in (\ref{ex:guo-dyn}) suggests a change out of the state of being a soldier, 
whereas \obj{le} in (\ref{ex:le-dyn}) conveys a change into the state of being a soldier \citep[149]{XiaoMcEnery2004}.

\ea
  \ea\label{ex:guo-dyn}
    \gll \objex{ta1} \objex{dang1-guo4} \objex{bing1}.\\
    he serve.as-\textsc{exp} soldier\\ \jambox{\citep[149]{XiaoMcEnery2004}}
    \glt `He once served as a soldier.'
  \ex\label{ex:le-dyn}
    \gll \objex{ta1} \objex{dang1-le} \objex{bing1}.\\
    he serve.as-\textsc{pfv} soldier\\ \jambox{\citep[149]{XiaoMcEnery2004}}
    \glt `He became a soldier.'
  \z
\z
It is clear that \obj{guo4} only indicates a change at the termination of a situation and cannot express the dynamicity within a situation.
Hence, it is incompatible with the semantics of the reduplication.

>>>>

\ea
  \ea\label{ex:le-dyn-look}
    \gll \objex{Wu2} \objex{Xu4mang2} \objex{kan4-le} \objex{zuo4-an4} \objex{shi2} \objex{liu2xia4} de \objex{jiao3yin1} \ldots\\ 
    Wu Xumang look-\textsc{pfv} commit-crime when leave \textsc{de} footprint\\ \jambox{\citep[158]{XiaoMcEnery2004}}
    \glt `Wu Xumang looked at the footprint left when the crime was committed.'
  \ex\label{ex:redup-dyn-look}
    \gll  \objex{Wu2} \objex{Xu4mang2} \obj{kan4-le-kan4} \objex{zuo4-an4} \objex{shi2} \objex{liu2xia4} de \objex{jiao3yin1} \ldots\\
    Wu Xumang look-\textsc{pfv}-look commit-crime when leave \textsc{de} footprint\\ \jambox{\citep[158]{XiaoMcEnery2004}}
    \glt `Wu Xumang looked a little bit at the footprint left when the crime was committed.'
  \z
\z

@@

As a reviewer points out, \citet[Sec.\,3.3]{Tham2013} considers  \obj{bing4} `be sick' to be a basic change of state (COS) verb, i.e. `become sick'.
In contrast, she uses \obj{xi3huan1} `like' and \obj{xiang1xin4} `believe' as examples of stative verbs.
\citet[680]{PeckEtAl2013} also list \obj{xi3huan1} `like' and \obj{xiang1xin4} `believe' as stative verbs.
For these two verbs, examples such as (\ref{ex:xihuan}) and (\ref{ex:xiangxin}) can be found.
\ea\label{ex:xihuan}%这车好型啊！！！。怎么了，我就喜欢喜欢 不行么~ (BCC)
\gll \objex{zhe4} \objex{che1} \objex{hao3} \objex{xing2} \objex{a}! \objex{zen3me} \objex{le}, \objex{wo3} \objex{jiu4} \obj{xi3huan1-xi3huan1} \objex{bu4xing2} \objex{me}?\\
this car so cool \textsc{ptc} what \textsc{ptc} I just like-like cannot \textsc{ptc}\\\jambox{(BCC)}
\glt `This car is so cool! What? I can't just like it for a little bit?'

\ex\label{ex:xiangxin}%谁让我来相信相信 (CCL)
\gll \objex{shei2} \objex{rang4} \objex{wo3} \objex{lai2} \obj{xiang1xin4-xiang1xin4}\\
who let I come believe-believe\\\jambox{(CCL)}
\glt `Who lets me believe him/her?'
\z

@@

\ea
  \ea\label{ex:cry}
    \gll \objex{wei4ci3}, \objex{Deng4} \objex{Li4jun1} \objex{shang1xin1de}  \objex{ku1-le} \objex{san1} \objex{tian1}.\\
    for.this Deng Lijun sadly  cry-\textsc{pfv} three day\\ \jambox{\citep[129]{XiaoMcEnery2004}}
    \glt `For this reason, Deng Lijun cried sadly for three days.'
  \ex\label{ex:know}
    \gll \objex{ta1} \objex{zhi1dao4-le} \objex{zhe4} \objex{shi4} de \objex{nei4-qing2}.\\
    he know-\textsc{pfv} this matter \textsc{de} inside-information\\ \jambox{\citep[130]{XiaoMcEnery2004}}
    \glt `He got to know the inside information on this matter.'
  \z
\z
%\obj{Le} is compatible with the reduplication, because its dynamicity can relate to not only the termination or instantiation of an event (a point of change), but also the process of the situation, just like that of the reduplication (see \sectref{sec:core-sem}).

In comparison, the experiential aspect marker \obj{guo} cannot co\hyp{}occur with a reduplicated verb, 
because its dynamicity attributes to an ``experiential change'' \citep[148]{XiaoMcEnery2004}, 
namely that a situation has been experienced historically and that ``the final state of the situation no longer obtains'' at the reference time \citep[144]{XiaoMcEnery2004}. 
Compare (\ref{ex:guo-dyn}) and (\ref{ex:le-dyn}), \obj{guo} in (\ref{ex:guo-dyn}) suggests a change out of the state of being a soldier, 
whereas \obj{le} in (\ref{ex:le-dyn}) conveys a change into the state of being a soldier \citep[149]{XiaoMcEnery2004}.

\ea
  \ea\label{ex:guo-dyn}
    \gll \objex{ta1} \objex{dang1-guo} \objex{bing1}.\\
    he serve.as-\textsc{exp} soldier\\ \jambox{\citep[149]{XiaoMcEnery2004}}
    \glt `He once served as a soldier.'
  \ex\label{ex:le-dyn}
    \gll \objex{ta1} \objex{dang1-le} \objex{bing1}.\\
    he serve.as-\textsc{pfv} soldier\\ \jambox{\citep[149]{XiaoMcEnery2004}}
    \glt `He became a soldier.'
  \z
\z
It is clear that \obj{guo} only indicates a change at the termination of a situation and cannot express the dynamicity within a situation.
Hence, it is incompatible with the semantics of the reduplication.
